---
title: "Modeling_Laura"
author: "Wanning Wang"
date: "4/15/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
pacman::p_load(
magrittr,
knitr,
sqldf,
lme4,
readxl,
lubridate,"grid","gridExtra","scales",
tidyverse,
)
```

Data Cleaning
-------------

### Read data
```{r}
# master <-read.delim("clipboard")

master <- read_xlsx("/Users/laura/OneDrive - Boston University/For Masanao Class - ROMS Full Data Set - March 19th, 2019 Upload.xlsx",sheet = 1)

claim <- read_xlsx("/Users/laura/OneDrive - Boston University/For Masanao Class - ROMS Full Data Set - March 19th, 2019 Upload.xlsx",sheet = 2)

# names(master)
check <- apply(master, 2, function(x) any(is.na(x)))

```

### clean colnames
```{r}
m1 <- as_tibble(master)
names(m1)<-str_replace_all(names(m1), c(" " = "_"))


colnames(m1)[7] <- "Sex"
colnames(m1)[11] <- "Outcome_index"
colnames(m1)[which(colnames(m1)=='Outcome_Minimal_Clinical_Important_Difference')] <- "Outcome_Mini"
colnames(m1)[which(colnames(m1)=='Chronic_Pain_(Yes/No)')]<-"Chronic_Pain"
colnames(m1)[which(colnames(m1)=='Length_Of_Stay_(days)')]<-"Length_Of_Stay"


## Injury_Date change to date
m1$Injury_Date[m1$Injury_Date=="5"] <-"NULL"
m1$Injury_Date <- as.numeric(m1$Injury_Date)
m1$Injury_Date <- as.Date(m1$Injury_Date, origin = "1899-12-30")

## Surgery_Date change to date
#str_detect(m1$Surgery_Date[1:10],"^0000-")

m1$Surgery_Date[which(m1$Surgery_Date=="Null")] <- "NULL"
m1$Surgery_Date[str_detect(m1$Surgery_Date,"^0000-")] <- "NULL"
m1$Surgery_Date <- as.numeric(m1$Surgery_Date)
m1$Surgery_Date <- as.Date(m1$Surgery_Date, origin = "1899-12-30")

## change Outcome mini score to numeric so that can compare
m1$Outcome_Mini <- as.numeric(m1$Outcome_Mini)

#which(colnames(m1)=='Sex_(1=male,_2=female)')
```

### Identify outcome 1/0 success or not on each vist and discharge base
```{r}
m1 %<>% mutate(Vis_Chg_Scr = ifelse(Outcome_index=="Bad", -1*(Visit_Outcome_Score - Admission_Outcome_Score),(Visit_Outcome_Score-Admission_Outcome_Score)))

m1 %<>% mutate(Vis_Chg_Scr = ifelse(Outcome_index=="Bad", -1*(Visit_Outcome_Score - Admission_Outcome_Score),(Visit_Outcome_Score-Admission_Outcome_Score)))

m1 %<>% mutate(Fin_Chg_Scr = ifelse(Outcome_index=="Bad",-1*(Outcome_Change_Scores),Outcome_Change_Scores))

m1 %<>% mutate(Vis_Suc = ifelse(Vis_Chg_Scr>=Outcome_Mini,1,0))
m1 %<>% mutate(Fin_Suc = ifelse(Fin_Chg_Scr>=Outcome_Mini,1,0))

```

### remove NA and abnormal values, convert text to upper case to category, remove duplicated values

```{r}
na.check <- function(x){

   na_count <-sapply(x, function(y) sum(length(which(is.na(y)))))
   na_count_df <- data.frame(na_count,col.n=seq(1:length(na_count)))
    return(View(na_count_df))
}
# na.check(m1)
## change the ROMS_ID 3957 one age value -967 ...-.-
m1$Age[!is.na(m1$Age) & (m1$Age<0)] <- m1$Age[which(!is.na(m1$Age) & (m1$Age<0))-1]

## tolower Outcome
m1 %<>% mutate(Outcome = toupper(Outcome))
m1 %<>% mutate(Age=round(Age))
m1 %<>% mutate(Body_Region=toupper(Body_Region))
m1 %<>% mutate(Start_Risk = toupper(Start_Risk))
m1 %<>% mutate(Chronic_Pain = toupper(Chronic_Pain))
m1$Chronic_Pain[m1$Chronic_Pain==1] <- "YES"
m1$Sex[m1$Sex==1]<- "MALE"
m1$Sex[m1$Sex==2]<- "FEMALE"

## remove duplicated m1
m1 <- m1 %>% distinct()

## m2 -- visit info

## m3 -- discharge info

#m3 <- m1 %>% select(ROMS_ID,Payer_Category,Age,Sex,Outcome,Outcome_Mini,Outcome_index,Body_Region,
                    #Surgical,Classification,Admission_Outcome_Score,Visits,Length_Of_Stay,Chronic_Pain,Start_Risk,Fin_Chg_Scr,F#in_Suc) %>% distinct()

m3 <- m1 %>% dplyr::select(-Clinic,-Visit_ID,-Outcome_Score_Range,-Pain_Minimal_Clinical_Important_Difference,-Pain_Score_Range,
                    -"Pain_-_Is_high_score_good_or_bad?",-Visit_Date,-Visit_Outcome_Score,-Discharge_Outcome_Score,-Discharge_Pain_Score,-Discharge_Date,
                    -Visit_Pain_Score,-Pain_Change_Scores,-Start_Back_Total,-Start_Back_Subscore,-Vis_Chg_Scr,-Vis_Suc) %>% distinct()

## Seperate Surgical and Conservative
m3.s <- m3 %>% filter(Surgical=="Surgical") 
m3.c <- m3 %>% filter(Surgical=="Conservative") %>% dplyr::select(-Surgery_Date) # --> focus

m3.c$Admission_Date <- as.Date(m3.c$Admission_Date)
m3.c$Injury_Date <- as.Date(m3.c$Injury_Date)

m3.c <- m3.c %>% mutate(week_dif = round(as.numeric(difftime(Admission_Date,Injury_Date,units = "weeks"))),days_dif = round(as.numeric(difftime(Admission_Date,Injury_Date,units = "days"))))


m3.c <- m3.c[-which( is.na(m3.c$week_dif) | (!is.na(m3.c$week_dif) & m3.c$week_dif<0)),]

quantile(m3.c$week_dif)
#  0%  25%  50%  75% 100% 
#  0    4   13   49 1252 

m3.c <- m3.c %>% mutate(vist_dif = ifelse(week_dif<=4,"<=4",
                                     ifelse(week_dif<=13,"5-13",
                                     ifelse(week_dif<=49,"14-49",">49"))))
```

## (updated model, delete visit info) Random Forest use m3.c matrix
```{r}
## Random Forest 

rf_df <- m3.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Injury_Date",-"Fin_Chg_Scr",-"week_dif",-"Outcome_Change_Scores",-"Length_Of_Stay",-"Visits") %>% na.omit()%>% mutate_if(is.character, as.factor)

rf_df$Fin_Suc <- as.factor(rf_df$Fin_Suc)

sapply(rf_df, class)

library(randomForest)
set.seed(2009)

train=sample(1:nrow(rf_df),round(nrow(rf_df)*0.8))
df_train = rf_df[train,]
df_test =  rf_df[-train,]

dim(df_train)
dim(df_test)


rf_fit1 = randomForest(Fin_Suc~.,data=df_train,importance=T)

rf_fit1

importance(rf_fit1)

## predict
fitted.results.rf1 <- predict(rf_fit1,newdata=df_test)

misClasificError <- mean(fitted.results.rf1 != df_test$Fin_Suc)
print(paste('Accuracy',1-misClasificError))
varImpPlot (rf_fit1)
plot(rf_fit1)
```

## New matrix take unique ROMS_ID as each observations
```{r}
## join claim to use Therapist
names(claim)<-str_replace_all(names(claim), c(" " = "_"))

join = "select m1.*, claim.Therapist from m1 left join claim on m1.Visit_ID=claim.Visit_ID"

m4 <- sqldf(join)

m4 <- m4 %>%  distinct()

m4 <- m4 %>% arrange(ROMS_ID, Visit_Date)

m4 <- m4 %>% filter(!is.na(Visit_ID))

m4 <- m4 %>% group_by(ROMS_ID) %>% mutate(id = row_number()) %>% ungroup()

## only take the first time patient visit in as  observations
m5 <- m4 %>% filter(id==1) #3346


m5 <- m5 %>% dplyr::select(-Clinic,-Visit_ID,-Outcome_Score_Range,-Pain_Minimal_Clinical_Important_Difference,-Pain_Score_Range,
                    -"Pain_-_Is_high_score_good_or_bad?",-Visit_Date,-Visit_Outcome_Score,-Discharge_Outcome_Score,-Discharge_Pain_Score,-Discharge_Date,
                    -Visit_Pain_Score,-Pain_Change_Scores,-Start_Back_Total,-Start_Back_Subscore,-Vis_Chg_Scr,-Vis_Suc) %>% distinct()

## Seperate Surgical and Conservative
m5.s <- m5 %>% filter(Surgical=="Surgical") 
m5.c <- m5 %>% filter(Surgical=="Conservative") %>% dplyr::select(-Surgery_Date) # --> focus

m5.c$Admission_Date <- as.Date(m5.c$Admission_Date)
m5.c$Injury_Date <- as.Date(m5.c$Injury_Date)

m5.c<- m5.c %>% mutate(week_dif = round(as.numeric(difftime(Admission_Date,Injury_Date,units = "weeks"))),day_dif = round(as.numeric(difftime(Admission_Date,Injury_Date,units = "days"))))
m5.c <- m5.c[-which( is.na(m5.c$day_dif) | (!is.na(m5.c$day_dif) & m5.c$day_dif<0)),]

quantile(m5.c$week_dif)
#   0%  25%  50%  75% 100% 
#   0    4   11   40 1252 

m5.c <- m5.c %>% mutate(vist_dif = ifelse(week_dif<=4,"<=4",
                                     ifelse(week_dif<=11,"5-11",
                                     ifelse(week_dif<=40,"12-40",">40"))))

quantile(m5.c$Admission_Outcome_Score)

```


## (updated model, delete visit info) Random Forest with therapist 
```{r}
rf_df_2 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Injury_Date",-"Fin_Chg_Scr",-"week_dif",-"Outcome_Change_Scores",-"id",-"Length_Of_Stay",-"Visits") %>% na.omit()%>% mutate_if(is.character, as.factor)

rf_df_2 $Fin_Suc <- as.factor(rf_df_2 $Fin_Suc)

sapply(rf_df_2, class)

#library(randomForest)
set.seed(2018)

train_2=sample(1:nrow(rf_df_2),round(nrow(rf_df_2)*0.80))
df2_train = rf_df_2[train_2,]
df2_test =  rf_df_2[-train_2,]

dim(df2_train)
dim(df2_test)

rf_fit2 = randomForest(Fin_Suc~.,data=df2_train,importance=T)

rf_fit2 

importance(rf_fit2)


## predict
fitted.results.rf <- predict(rf_fit2,newdata=df2_test)

misClasificError <- mean(fitted.results.rf != df2_test$Fin_Suc)
print(paste('Accuracy',1-misClasificError))
varImpPlot (rf_fit2)
```

##(updated model, delete visit info) Random Forest without therapist [in W4 slide]
```{r}
rf_df_3 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Fin_Chg_Scr",-"vist_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Visits",-"day_dif") %>% na.omit()%>% mutate_if(is.character, as.factor)

rf_df_3 $Fin_Suc <- as.factor(rf_df_3 $Fin_Suc)

sapply(rf_df_3, class)

#library(randomForest)
set.seed(678302)

train_3=sample(1:nrow(rf_df_3),round(nrow(rf_df_3)*0.80))
df3_train=  rf_df_3 [train_3,]
df3_test =  rf_df_3 [-train_3,]

dim(df3_train)
dim(df3_test)
```


```{r}
rf_fit3 = randomForest(Fin_Suc~.,data=df3_train,importance=T,
nodesize=10,mtry=5,ntree=450)

rf_fit3

importance(rf_fit3)


## predict
fitted.results.rf2 <- predict(rf_fit3,newdata=df3_test)

misClasificError <- mean(fitted.results.rf2 != df3_test$Fin_Suc)
print(paste('Accuracy',1-misClasificError))
p <- varImpPlot (rf_fit3)



var_importance <- data_frame(variable=setdiff(colnames(df3_train), "Fin_Suc"),
                             importance=as.vector(importance(rf_fit3,type = 2)))
var_importance <- arrange(var_importance, desc(importance))
var_importance$variable <- factor(var_importance$variable, levels=var_importance$variable)

p <- ggplot(var_importance, aes(x=variable, weight=importance, fill=variable))
p <- p + geom_bar() + ggtitle("Variable Importance from Random Forest Fit")
p <- p + xlab("Predictor") + ylab("Variable Importance (Mean Decrease in Gini Index)")
p <- p + scale_fill_discrete(name="Predictor Name")
p + theme(axis.text.x=element_blank(),
          axis.text.y=element_text(size=12),
          axis.title=element_text(size=12),
          plot.title=element_text(size=12,face="bold"),
          legend.title=element_text(size=12),
          legend.text=element_text(size=12))



#plot(rf_fit3)
```


```{r}
```


## (updated model, delete visit info) Random Forest in different outcome categories [in W4 slide]
```{r}
rf_df_5 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Injury_Date",-"Fin_Chg_Scr",-"vist_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Outcome_index",-"Visits") %>% na.omit()%>% mutate_if(is.character, as.factor)


rf_df_5 $Fin_Suc <- as.factor(rf_df_5 $Fin_Suc)
out.cat <- unique(rf_df_5$Outcome)
br.cat <- unique(rf_df_5$Body_Region)
#LOWER EXTREMITY FUNC SCALE 

df.out <- vector("list",length(out.cat))
out.train <- vector("list",length(out.cat))
out.test <- vector("list",length(out.cat))
rf.fit.out.s <- vector("list",length(out.cat))
fitted.results.rf.out <- vector("list",length(out.cat))

accuracy <- rep(NA,length(out.cat))

j = 1
for (i in out.cat){
  set.seed(2018)
  df.out[[j]]=rf_df_5 %>% filter(Outcome==i)
  train.out=sample(1:nrow(df.out[[j]]),round(nrow(df.out[[j]])*0.80))
  out.train1 = df.out[[j]]
  out.train = out.train1[train.out,]
  out.test1 = df.out[[j]]
  out.test =out.test1[-train.out,]
  rf.fit.out = randomForest(Fin_Suc~.,data= out.train,importance=T)
  rf.fit.out.s[[j]]=rf.fit.out
  fitted.results.rf.out <- predict(rf.fit.out,newdata=out.test)
  misClasificError <- mean(fitted.results.rf.out!=out.test$Fin_Suc)
  #print(paste(i,'Accuracy',1-misClasificError))
  accuracy[j]= paste(i,'Accuracy',1-misClasificError)
  j = j+1
}

accuracy
rf.fit.out.s
```

```{r}
rf_df_5 %>% group_by(Outcome) %>% summarize(cases=n()) %>% arrange(desc(cases))
```


```{r}
accuracy2 <- rep(NA,length(br.cat))

j = 1
for (i in br.cat){
  set.seed(2019)
  df.out[[j]]=rf_df_5 %>% filter(Body_Region==i)
  train.out=sample(1:nrow(df.out[[j]]),round(nrow(df.out[[j]])*0.80))
  out.train1 = df.out[[j]]
  out.train = out.train1[train.out,]
  out.test1 = df.out[[j]]
  out.test =out.test1[-train.out,]
  rf.fit.out = randomForest(Fin_Suc~.,data= out.train,importance=T)
  rf.fit.out.s[[j]]=rf.fit.out
  fitted.results.rf.out <- predict(rf.fit.out,newdata=out.test)
  misClasificError <- mean(fitted.results.rf.out!=out.test$Fin_Suc)
  print(paste(i,'Accuracy',1-misClasificError))
  accuracy2[j]= paste(i,'Accuracy',1-misClasificError)
  j = j+1
}

accuracy2

rf.fit.out.s


rf.fit.out.s
```





## logistic regression use m5.c matrix
```{r}
library(MASS)

logistic_fit1 <- glm(Fin_Suc~.,data=df2_train,family=binomial(link="logit"))

summary(logistic_fit1)

fitted.results.l1 <- predict(logistic_fit1,newdata=df2_test,type='response')
fitted.results.l1 <- ifelse(fitted.results.l1 > 0.5,1,0)
misClasificError <- mean(fitted.results.l1 != df2_test$Fin_Suc)
print(paste('Accuracy',1-misClasificError))

```


## prepare data for treeshap
```{r}
rf_df_3 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Injury_Date",-"Fin_Chg_Scr",-"vist_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Visits",-"Outcome_index") %>% na.omit()%>% mutate_if(is.character, as.factor)

rf_df_3 $Fin_Suc <- as.factor(rf_df_3 $Fin_Suc)


## data with week_dif
rf.df1 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Fin_Chg_Scr",-"Outcome_Change_Scores",-"id",-"Length_Of_Stay",-"Visits",-"Therapist") %>% na.omit()%>% mutate_if(is.character, as.factor)
rf.df1 $Fin_Suc <- as.factor(rf.df1 $Fin_Suc)

## data with vist_dif
rf.df2 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Fin_Chg_Scr",-"week_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Visits") %>% na.omit()%>% mutate_if(is.character, as.factor)
rf.df2 $Fin_Suc <- as.factor(rf.df2 $Fin_Suc)


## data with vist_dif
rf.df2 <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Fin_Chg_Scr",-"week_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Visits") %>% na.omit()%>% mutate_if(is.character, as.factor)
rf.df2 $Fin_Suc <- as.factor(rf.df2 $Fin_Suc)

## write out data
write.csv(rf.df1,"/Users/laura/OneDrive - Boston University/rf1.csv",row.names=FALSE)
write.csv(rf.df2,"/Users/laura/OneDrive - Boston University/rf2.csv",row.names=FALSE)
write.csv(rf.df3,"/Users/laura/OneDrive - Boston University/rf2.csv",row.names=FALSE)


ggplot(rf.df1) + geom_bar(aes(x=Start_Risk,fill=Fin_Suc)) #+ facet_wrap(~Outcome)


```


## logistic
```{r}
lg_df <- m5.c %>% dplyr::select(-"ROMS_ID",-"Payer",-"Outcome_Mini",-"Surgical",-"Classification",-"Admission_Date",-"Injury_Date",-"Injury_Date",-"Fin_Chg_Scr",-"vist_dif",-"Outcome_Change_Scores",-"id",-"Therapist",-"Length_Of_Stay",-"Visits",-"day_dif",-'Payer_Category') %>% na.omit()%>% mutate_if(is.character, as.factor)

lg_df $Fin_Suc <- as.factor(lg_df $Fin_Suc)

sapply(lg_df, class)

#library(randomForest)
set.seed(678302)

lg=sample(1:nrow(lg_df),round(nrow(lg_df)*0.80))
lg_train=  lg_df [lg,]
lg_test =  lg_df [-lg,]


library(MASS)

logistic_fit1 <- glm(Fin_Suc~.,data=lg_train,family=binomial(link="logit"))

summary(logistic_fit1)

fitted.results.l1 <- predict(logistic_fit1,newdata=lg_test,type='response')
fitted.results.l1 <- ifelse(fitted.results.l1 > 0.5,1,0)
misClasificError <- mean(fitted.results.l1 != lg_test$Fin_Suc)
print(paste('Accuracy',1-misClasificError))
```

